Comparing probabilistic predictive models applied to football

### 2015-07-04 template.tex
com sugestoes do ernesto na secao 3

	Formally, let $X$ be a random variable indicating
	the result of a given match. Hence, $X$ assumes a value in $\{1,2,3\}$. Moreover, let $P$ denote one's probabilistic prediction
	about $X$, i.e., $P$ is such that $\sum_{x=1}^3P(X=x)=1$ and
    $P(X=x)\geq0$ for all $x$ in $\{1,2,3\}$.
	A scoring rule is a function 
	that assigns a real number (score) $S(x,P)$ to each combination $(x,P)$, where $x$ denotes each possible outcome of the match.
	For a given $x$ in $\{1,2,3\}$, the score  $S(x,P)$ is minimum when $P$ is degenerate distribution on $x$, i.e., when $P(X=x)=1$.  
	Thus, the best score (i.e., the score with the smallest value) is achieved when the probabilistic prediction for the outcome of the game is a degenerate  distribution
  on the observed outcome of that match, i.e., a perfect prediction.\footnote{A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good forecasts.}
	%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

	Although many functions satisfy the criteria for a scoring rule specified above, 
	not all scoring rules encourage honesty and accuracy in assessing numerical values for their predictions.
	Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}, which we describe in the sequence.

	Say a model's probabilistic prediction for $X$ is $P^*.$ A proper scoring rule is a scoring rule $S$ such that the probabilistic prediction $P$ that minimizes 
	$$E_{P^*}[S(X,P)]=\sum_{x=1}^3 S(x,P)P^*(X=x)$$
	 is $P^*$  itself.
	In other words, if one imagines 
	a penalty $S(x,P)$ is received 
	when $x$ happens then, 
	if $S$ is a proper scoring rule, the lowest expected penalty is obtained by reporting the model's  uncertainty about $X$, $P^*$.
	The use of a proper scoring rule thus encourages the forecaster to use $P^*$
	as his probabilistic prediction for the outcome of the game, $X$, rather than some other quantity. 
	%Moreover, this method automatically leads to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each model.

@@

	Let $P$ be a probabilistic prediction  for $X$.
	The Brier score for a given outcome $x\in\{1,2,3\}$ is given by
	$$S(x,P)= \sum_{i=1}^3\mathbb{I}(x=i)(1- P(X=i))^2+\sum_{i=1}^3\mathbb{I}(x\neq i)P^2(X=i),$$
	where $\mathbb{I}$ is the indicator function.

@@

	The Brier score for the probabilistic prediction
	$$P(X=1)=0.25, P(X=2)=0.35, P(X=3)=0.40,$$ 
	assuming the home team loses, is therefore given by $d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. 
	On the other hand, the prediction $P(X=3)=1$ achieves score  zero, the minimum for this rule.

	It is useful to consider the score of what we will call {\it trivial prediction}:
	$$P(X=1)=1/3, P(X=2)=1/3, P(X=3)=1/3.$$
	This assessment will produce a Brier score of $2/3$, no matter what is the final result of the match, providing, thus, a threshold that a good model should consistently beat, meaning, for the Brier score, that the scores of its predictions should be smaller than $0.667$.

@@

	The logarithmic  score is given by
	$$S(x,P)=- \sum_{i=1}^3\mathbb{I}(x=i)\ln(P(X=i)),$$

@@

	The logarithmic score for the prediction 
		$$P(X=1)=0.25, P(X=2)=0.35, P(X=3)=0.40,$$ 
		 when the home team loses is therefore 
	$-\ln(0.4)\approx 0.91$. 
	On the other hand, the prediction $P(X=3)=1$ achieves score zero, once again the minimum of this rule.
	Moreover, for the logarithmic score, the trivial prediction gives a score of approximately $1.098$.

@@

	The spherical score is given by
	$$S(x,P)=- \frac{1}{\sqrt{\sum_{i=1}^3 P^2(X=i)}}\sum_{i=1}^3\mathbb{I}(x=x_i)P(X=i),$$

@@

	The spherical score for the prediction 
		$$P(X=1)=0.25, P(X=2)=0.35, P(X=3)=0.40,$$ 
		 assuming the home team loses, is given by
	$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$.
	On the other hand, the prediction $P(X=3)=1$ achieves score  $-1$ instead and, for this rule, the trivial prediction results in a score of approximately $-0.577$.

@@

	The first one is the proportion of errors made by the model or assessor. This is simply the proportion of mistakes made when considering the highest probability assessment.
	More precisely, the proportion of errors of a given probabilistic prediction $P$ is defined by
	$$\frac{1}{n}\sum_{j=1}^n \mathbb{I}\left(X_j \neq \arg \max_{x \in \{1,2,3\}} P(X_j=x)\right),$$
	where $X_j$ is the outcome of the $j$-th match.

@@

	Because we typically do not have several predictions with the same assigned probability $p$, we obtain a plot by smoothing (i.e., regressing) the indicator function of whether a given result happened as a function of the probability assigned for that result.
	That is, we estimate the probability that a event occurs given its assigned probability.
	%, that is, we regress  
	%$$\{ \mathbb{I}_{\{i\}}(X_j) \}_{j=1,\ldots,n;i=1,2,3} \mbox{ over } \{P_X(x_j)\}_{j=1,\ldots,n;i=1,2,3}$$
	% for each event $A_i$, i.e., each possible outcome of each match
	%that occurred.
	The smoothing is done via smoothing splines \cite{wahba}, with tuning parameters chosen by cross-validation.

>>>>

	Formally, let $X$ be a random variable
	taking values in $\mathcal{X}=\{1,2,3\}$ indicating
	the outcome of the match, with 1 standing for home win, 2 for draw and 3 for away win.  Moreover, let $P=(P_1,P_2,P_3)$ denote one's probabilistic prediction
	about $X$, i.e., $P$ lies in the 2-simplex set $\Delta_2=\{(p_1,p_2,p_3):p_1+p_2+p_3=1, \ p_1,p_2,p_3\geq0\}$ (see Figure
	\ref{fig:simplex}). 	
	A scoring rule is a function 
	that assigns a real number (score) $S(x,P)$ to each $x \in \mathcal{X}$
	and $P \in \Delta_2$ 
	such that 
	for any given $x$ in $\mathcal{X}$, the score  $S(x,P)$ is minimized when $P$ is 
	$(1,0,0)$, $(0,1,0)$ or $(0,0,1)$ depending if $x$ is 1, 2 or 3, respectively.
	The score $S(x,P)$ can be thought as
	a penalty to be paid when one assigns the 
	probabilistic prediction $P$ and outcome
	$x$ occurs. Also, the ``best"\footnote{A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good forecasts.} possible score (i.e., the smallest score value) is achieved when the probabilistic prediction for the outcome of the game is perfect.
	%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

	Although many functions can satisfy the above scoring rule definition, not all of them encourage honesty and accuracy when assigning a prediction to an event. Those that do enable a fair probabilistic assignment are named \emph{proper scoring rules} \cite{lad}, which we describe in the sequence.

	Say a model's probabilistic prediction for $X$ is $P^*=(P_1^*,P_2^*,P_3^*).$ A proper scoring rule $S$ is a scoring rule  such that the mean score value 
	$$E_{P^*}[S(X,P)]=\sum_{x=1}^3 S(x,P)P^*_x$$
	 is minimized when $P=P^*$.
	In other words, if one announces
	$P$ as his probabilistic prediction
	and $S$ as a scoring rule, the lowest
	expected penalty is obtained by reporting the correct model's  uncertainty about $X$, $P^*$.
	Thus, the use of a proper scoring rule encourages the forecaster to announce $P^*$ (the correct one)
	as his probabilistic prediction  rather than some other quantity. 
	%Moreover, this method automatically leads to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each model.

@@

	Let $P=(P_1,P_2,P_3)$ be a probabilistic prediction  for $X$.
	The Brier score for a given outcome $x\in\{1,2,3\}$ is given by
	$$S(x,P)= \sum_{i=1}^3\mathbb{I}(x=i)(1- P_i)^2+\sum_{i=1}^3\mathbb{I}(x\neq i)P^2_i,$$
	where $\mathbb{I}$ is the indicator function.

@@

	The Brier score for the probabilistic prediction $P=(0.25,0.35,0.40)$
	assuming the home team loses, is therefore given by $d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. 
	On the other hand, the prediction $P=(0,0,1)$ achieves score  zero, the minimum for this rule.

	It is useful to consider the score of what we will call {\it trivial prediction}:
	$P=(1/3,1/3,1/3)$.
	This assessment will produce a Brier score of $2/3$, no matter what is the final result of the match, providing, thus, a threshold that a good model should consistently beat, meaning, for the Brier score, that the scores of its predictions should be smaller than $0.667$.

@@

	The logarithmic  score is given by
	$$S(x,P)=- \sum_{i=1}^3\mathbb{I}(x=i)\ln(P_i),$$

@@

	The logarithmic score for the prediction 
	$P=(0.25,0.35,0.40)$
		 when the home team loses is therefore 
	$-\ln(0.4)\approx 0.91$. 
	On the other hand, the prediction $P=(0,0,1)$ achieves score zero, once again the minimum of this rule.
	Moreover, for the logarithmic score, the trivial prediction gives a score of approximately $1.098$.

@@

	The spherical score is given by
	$$S(x,P)=- \frac{1}{\sqrt{\sum_{i=1}^3 P^2_i}}\sum_{i=1}^3\mathbb{I}(x=x_i)P_i,$$

@@

	The spherical score for the prediction 
		$P=(0.25,0.35,0.40)$ assuming the home team loses, is given by
	$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$.
	On the other hand, the prediction $P=(0,0,1)$ achieves score  $-1$ instead and, for this rule, the trivial prediction results in a score of approximately $-0.577$.

@@

	The first one is the proportion of errors made by the model or assessor. This is simply the proportion of mistakes made when considering the highest probability assessment.
	More precisely, the proportion of errors of a a sequence of probabilistic predictions for $n$ games, $P^{(1)},\ldots,P^{(n)}$
	with
		$P^{(j)}=(P^{(j)}_1,P^{(j)}_2,P^{(j)}_3)$, is defined by
	$$\frac{1}{n}\sum_{j=1}^n \mathbb{I}\left(X_j \neq \arg \max_{x \in \{1,2,3\}} P^{(j)}_x\right),$$
	where $X_j$ is the outcome of the $j$-th match.

@@

	Because we typically do not have several predictions with the same assigned probability $p$, we obtain a plot by smoothing (i.e., regressing) the indicator function of whether a given result happened as a function of the probability assigned for that result.
	That is, we estimate the probability that an event occurs given its assigned probability.
	%, that is, we regress  
	%$$\{ \mathbb{I}_{\{i\}}(X_j) \}_{j=1,\ldots,n;i=1,2,3} \mbox{ over } \{P_X(x_j)\}_{j=1,\ldots,n;i=1,2,3}$$
	% for each event $A_i$, i.e., each possible outcome of each match
	%that occurred.
	The smoothing is done via smoothing splines \cite{wahba}, with tuning parameters chosen by cross-validation.
